Enhanced Material Point Method to face dynamic problems: Local-maximum entropy approximation and explicit predictor-corrector scheme\\ \PNA{On the dynamic assessment of the Local-Maximum Entropy Material Point Method through an Explicit Predictor-Corrector Scheme}

### 2020-04-27 LME-MPM.tex
Quedan_conclusiones

The following test was proposed to measure proof the ability of this
interpolation technique to deal with grid crossing instabilities. It
consists in the simulation of a square block of soil incrementally
loaded by a body force. Details of the problem are sketched in figure~\ref{fig:block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.7\hsize}{!}{
    \input{Figures/Block.tex}}
  \caption{Geometrical description of a soil block }.
  \label{fig:block}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
This test was taken from PhD thesis of Andersen (2009)\cite{thesis_Andersen_2009}. The
elastic parameters consider for this test are: 
\begin{itemize} 
\item  Initial density : $6\cdot 10^3\ kg/m^3$
\item  Poisson ratio : $0$
\item  Elastic modulus : $5\ MPa$
\end{itemize}
The gravity force is a apply as an external force according to the
equations \eqref{eq:particle_body_forces},
\eqref{eq:nodal_external_forces}. Using a total time period of T (20
seconds) to apply the gravity, it is increased from 0 to 9.81$m/s$
with a sinus function until T/2 and then maintained constant until T
in order to arrive to a state of equilibrium, 
\begin{equation}
  \label{eq:gravity-load-block}
 \mathbf{g}(t) = \left\{
    \begin{array}{ll}
      0.5 \mathbf{g} (\sin(\frac{2t \pi}{T} - \frac{\pi}{2})+1)  & \mbox{if } t \leq T/2 \\
      \mathbf{g} & \mbox{if } t > T/2
    \end{array}
  \right.
\end{equation}
In order to get a stable solution, time step was conducted by a
Courant number of 0.1. On the other hand, the explicit
predictor-corrector scheme is here employed looking forward getting better results. For the
initial spatial discretization four particles per cell
($\Delta x = 2\ m$) were adopted. The initial layout of particles inside of the
cell changes according to the approximation technique adopted. For the
bi-linear shape functions and the \acrshort{lme} approximants, the initial
position corresponds to the location of the gauss-points in a standard
quadratic finite element. For the \acrshort{ugimp} shape function the initial
position of each particle is located in the center of each voxel, due
to the fact that in the initial situation, the voxel domain should not
overlap each others.

Figure~\ref{fig:Block-LME3} shows the evolution of the vertical stress
during the loading process. The result is physically realistic as
stress increments linearly from the top to the bottom of the specimen,
and the value of the vertical stress in a material point located in
the bottom of the specimen oscillates centered in $5.2 MPa$, which is
the analytic value given by $\sigma_{yy} = \rho g h_y$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
  \centering
  \subfigure[t = 0 seconds.]{
    \includegraphics[width=0.3\textwidth]{Figures/Block_LME3_PCE_a_t0}
  }
  \subfigure[t = 5 seconds.]{
    \includegraphics[width=0.3\textwidth]{Figures/Block_LME3_PCE_b_t025}
  }
  \subfigure[t = 20 seconds]{
    \includegraphics[width=0.3\textwidth]{Figures/Block_LME3_PCE_c_t1}
  }
  \caption{Vertical normal stress and position of material points
    during the loading process for a soft soil ($E = 5\ MPa$, $\rho_0
    = 6\cdot 10^3\ kg/m^3$). Numerical parameters considered for the
    simulation are : Local \textit{max-ent} shape function $\gamma =3$
    and explicit PC scheme with CFL 0.1.}
  \label{fig:Block-LME3}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Figure~\ref{fig:vertical-displacement-block} shows the vertical
displacement evolution of a point in the free surface of the
block. This figure shows how simulations performed with a bi-linear
interpolation technique (Q4) turns out to be unstable during
cell-crossing and consequently fails. The \acrshort{ugimp} simulation
is more stable than the one performed by the Q4. Despite this is still
unstable and could trigger severe oscillations in simulation with
non-linear  materials. The \acrshort{lme} simulation was performed
using two kinds of shape functions, one with a low value of the
dimensionless parameter, $\gamma = 0.8$, and other with a larger value
of it, $\gamma = 3.0$. Notice that the results are both stable, but
the larger values of $\gamma$ give us a very stable solution. This is
due to the fact that with larger value of $\gamma$, the shape
functions behaves in a similar way to the \acrshort{fem}, which performs very
accurate in those cases with a reasonable mesh distorsion, and with a
lower value it behaves in a similar way to the \acrshort{ugimp}. This behaviour
was noticed previously by \cite{Arroyo2006}, were authors highlight
how by adjusting the spatial variation of $\beta(\vect{x})$, it is
possible to select regions of the domain of analysis which are treated
by finite elements and regions that are treated in the style of
meshfree methods, with seamless transitions between those regions. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Block_CFL_01_Comparative}
  }
  \caption{Comparative of the vertical displacement evolution in a
    point located in the free surface employing different
    interpolation schemes and numerical techniques.} 
  \label{fig:vertical-displacement-block}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

The following test was proposed to \PNA{validate} \cancel{measure proof} the ability of the proposed
interpolation technique to deal with grid crossing instabilities. It
consists of the simulation of a square block of soil incrementally
loaded by a body force. Details of the problem are sketched in figure~\ref{fig:block}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{0.7\hsize}{!}{
    \input{Figures/Block.tex}}
  \caption{Geometrical description of a soil block }.
  \label{fig:block}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
This test was \cancel{taken from PhD thesis of Andersen} \PNA{previously proposed by Andersen} (2009)\cite{thesis_Andersen_2009}. The
elastic parameters considered for this test are: 
\begin{itemize} 
\item  Initial density : $6\cdot 10^3\ kg/m^3$
\item  Poisson ratio : $0$
\item  Elastic modulus : $5\ MPa$
\end{itemize}
The gravity force is applied as an external force according to the
equations \eqref{eq:particle_body_forces} and
\eqref{eq:nodal_external_forces}. Using a total time period of T (20
seconds) to apply the gravity, it is increased from 0 to 9.81$m/s$
with a sinusoidal function until T/2 and then maintained constant until T
in order to reach the equilibrium state:
\begin{equation}
  \label{eq:gravity-load-block}
 \mathbf{g}(t) = \left\{
    \begin{array}{ll}
      0.5 \mathbf{g} (\sin(\frac{2t \pi}{T} - \frac{\pi}{2})+1)  & \mbox{if } t \leq T/2 \\
      \mathbf{g} & \mbox{if } t > T/2
    \end{array}
  \right.
\end{equation}
In order to get a stable solution, time step was conducted by a
Courant number of 0.1. On the other hand, the explicit
predictor-corrector scheme is here employed looking forward to getting better results. For the
initial spatial discretization four particles per cell
($\Delta x = 2\ m$) were adopted. The initial layout of particles inside of the
cell changes according to the approximation technique adopted. For the
bi-linear shape functions and the \acrshort{lme} approximants, the initial
position corresponds to the location of the gauss-points in a standard
quadratic finite element. For the \acrshort{ugimp} shape function the initial
position of each particle is located in the center of each voxel, due
to the fact that in the initial situation, the voxel domain should not
overlap each others.

Figure~\ref{fig:Block-LME3} shows the evolution of the vertical stress
during the loading process. The result is physically realistic as
the stress increases linearly from the top to the bottom of the specimen,
and the value of the vertical stress in a material point located in
the bottom of the specimen oscillates around $5.2 MPa$, which is
the analytic value given by $\sigma_{yy} = \rho g h_y$.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
  \centering
  \subfigure[t = 0 seconds.]{
    \includegraphics[width=0.3\textwidth]{Figures/Block_LME3_PCE_a_t0}
  }
  \subfigure[t = 5 seconds.]{
    \includegraphics[width=0.3\textwidth]{Figures/Block_LME3_PCE_b_t025}
  }
  \subfigure[t = 20 seconds]{
    \includegraphics[width=0.3\textwidth]{Figures/Block_LME3_PCE_c_t1}
  }
  \caption{Vertical normal stress and position of material points
    during the loading process for a soft soil ($E = 5\ MPa$, $\rho_0
    = 6\cdot 10^3\ kg/m^3$). Numerical parameters considered for the
    simulation are : Local \textit{max-ent} shape function $\gamma =3$
    and explicit PC scheme with CFL 0.1.}
  \label{fig:Block-LME3}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Figure~\ref{fig:vertical-displacement-block} shows the vertical
displacement evolution of a point in the free surface of the
block. This figure shows how simulations performed with a bi-linear
interpolation technique (Q4) turns out to be unstable during
cell-crossing and consequently fail. The \acrshort{ugimp} simulation
is more stable than the one performed by the Q4, \PNA{although} \cancel{Despite this} is still
unstable and could trigger severe oscillations in simulation with
non-linear  materials. The \acrshort{lme} simulation was performed
using two kinds of shape functions, one with a low value of the
dimensionless parameter, $\gamma = 0.8$, and other with a larger value
of it, $\gamma = 3.0$. Notice that the results are both stable, but
the larger values of $\gamma$ give us a very stable solution. This is
due to the fact that with larger value of $\gamma$, the shape
functions behaves in a similar way to the \acrshort{fem}, which performs very
accurate in those cases with a reasonable mesh distorsion, and with a
lower value it behaves in a similar way to the \acrshort{ugimp}. This behaviour
was noticed previously by \cite{Arroyo2006}, were authors highlighted
how, by adjusting the spatial variation of $\beta(\vect{x})$, it is
possible to select regions of the domain of analysis which are treated
by finite elements and regions that are treated in the style of
meshfree methods, with seamless transitions between those regions. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}\sidecaption
  \centering
  \resizebox{\hsize}{!}{
    \includegraphics[width=\textwidth]{./Figures/Block_CFL_01_Comparative}
  }
  \caption{Comparative of the vertical displacement evolution in a
    point located in the free surface employing different
    interpolation schemes and numerical techniques.} 
  \label{fig:vertical-displacement-block}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
